### 2016-01-07 papers/CAFe/HIPS-paper/coarray-extensions.tex
Added text about remote execution.

The final CAFe concept introduced is that of remote execution.  As discussed, in Fortran 2008
all images (similar to MPI ranks) execute the same program.  The only way for an image
to execute something different is to conditionally execute a block of code based on explicitly
checking that an executing image's rank (\texttt{this\_image()}) meets some established criteria.
There exists no mechanism for one image to execute a block of code or a procedure on another
image.

CAFe allows images to execute tasks on hosted subimages using standard procedure calls.
For example,

>>>>

The final CAFe concept introduced is that of remote execution.  As discussed above, in
Fortran all CAF images (similar to MPI ranks) execute the same program.  The only way for
an image to execute something different is to conditionally execute a block of code based
on explicitly checking that an executing image's rank (\texttt{this\_image()}) meets some
established criterion.  There currently exists no mechanism for one image to execute a
block of code or a procedure on another image.

CAFe allows images to execute tasks on hosted subimages using standard procedure calls.
For example,
\begin{verbatim}
  call relax(U[device])  [[device]]
\end{verbatim}
executes the subroutine relax, on the subimage device, using array \texttt{U} \emph{located}
on the same subimage.  This follows the same pattern as introduced for remote coarray allocation,
whereby the double square brackets notation indicates \emph{where} the function will execute.
Funtions to be executed on a subimage must be pure procedures (i.e., declared with the Fortran
keyword \texttt{pure}) and scalar formal parameters must be passed by value (i.e., declared with the
\texttt{value} attribute).

CAFe expands on CAF execution semantics by providing a task-based mechanism, where tasks
are defined in terms of pure Fortran procedures.  Since the procedures must be declared
as pure, they cannot perform any I/O or call any impure procedures; nor can they initiate
any communication between the subimage and the hosting image.

CAFe subimage tasks can also be defined by the body of a Fortran do concurrent construct
and executed on a subimage.  For example, the code segment,
\begin{verbatim}
do concurrent(i=1:N-1)  [[device]]
 begin block
   real :: w = 0.67
   S(i) = (1-w)*T(i) + w*(T(i-1)+T(i+1)/2)
 end block
end do
\end{verbatim}
will execute all of the code within the \texttt{do concurrent} construct on device.  In
Fortran, the execution semantics of a \texttt{do concurrent} loop is that the loop
iterates can be executed concurrently in any order.

Note the similarities between this example and the OpenMP pragma \texttt{!\$omp parallel do,
private(w)}.  The use of the Fortran block construct to declare private
variables allows a natural and clear way to distinguish between variables that are private
to a ``thread'' (although in the example above, the ``private'' distinction for the
variable \texttt{w} is unnecessary because it is a constant for all loop iterations).
(NOTE FOR CRAIG: need to read the standard to see if a block construct can be used in this
way. ALSO, there is a lot more work here to think about how do concurrent merges with
OpenMP and OpenACC with regards to synchronization, barriers, teams, atomics,
, critical sections, ...)
